---
title: 'Denecke et. al 2017: Describing the Role of Drosophila melanogaster ABCB Transporters
  in Insecticide Biology Using CRISPR-Cas9 Knockouts'
output:
  html_document: default
  html_notebook: default
---

```{css, include=FALSE}
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F,message=F)
dir.create("~/test_ABCB")
knitr::opts_knit$set(root.dir = "~/test_ABCB")
```



# Prior To Beginning

Welcome to the R markdown walk through for the paper *Welcome to the R markdown companion to the recent publication Describing the Role of Drosophila melanogaster ABCB Transporters in Insecticide Biology Using CRISPR-Cas9 Knockouts* (Denecke et. al 2017). This document is designed to guide the reader through every step of the analysis accomplished in the paper and allow for the reproducibility of the study.


Before beginning a few housekeeping issues must be addressed. Firstly, all code demonstrated here was optimized on a linux machine but has been tested on a mac and a PC. While it should work the particular specifications of the environment in which it was written are listed below.


```{r Display Version}
version
```


Next, the analysis will require several packages. These packages can be installed and loaded into the R environment with the following chunk of code.

```{r Install and Download Packages}
list.of.packages <- c("devtools","plotrix")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if(!"insect.toxicology" %in% installed.packages()) devtools::install_github("shanedenecke/insect.toxicology")

library(insect.toxicology)
library(plotrix)
```

The data will need to be downloaded from the Mendeley repository. This can be done with the follwoing code.

```{r Download Data}
#download.file("https://data.mendeley.com/datasets/92sx2h3kfs/1/files/10fd8799-4c51-4efc-98a4-80a438102d30/Fig2_Mdr65KO_Lc50.csv?dl=1",destfile="~/test_ABCB/Denecke_ABCB.zip")
#unzip("~/test_ABCB/Denecke_ABCB_.zip")
```

Lastly, some custom functions that are used during the analysis can be loaded now. 
```{r Basic Functions}

firstup <- function(x) {
      substr(x, 1, 1) <- toupper(substr(x, 1, 1))
      x
}    


multi.gsub <- function(vector,find,replace){
      for (i in 1:length(find)){
            vector <- gsub(find[i],replace[i],vector)
      }
      return(vector)
} 
```


```{r Read in Data for Figure 1}
abcb.combined <-read.csv("Fig1_ABCB_Combined.csv")
```

```{r Import Figure 1}
abcb.abbots <- abbots.correction.table(abcb.combined,control="wxac",write=F) %>%
      filter(dose!=0)

```



```{r Format figure 1}

index <- abcb.abbots %>% select(genotype,pesticide) %>% unique() 
summary.matrix <- matrix(data=NA,nrow=length(unique(index$pesticide)),ncol=length(unique(index$genotype)),dimnames=list(unique(index$pesticide),unique(index$genotype)))
control="wxac"
for(i in 1:nrow(index)){
      test.genotype <- as.character(index[i,1])
      test.pesticide <- as.character(index[i,2])
      sub <- subset(abcb.abbots, (genotype==test.genotype | genotype==control) & pesticide==test.pesticide)
      emvec <- c()
      for(d in unique(sub$dose)){
            sub2 <- subset(sub,dose==d)
            if(sub2$lcl[which(sub2$genotype=="wxac")] > sub2$ucl[which(sub2$genotype==test.genotype)]){
                  emvec <- c(emvec,"sus")
            }else if (sub2$ucl[which(sub2$genotype=="wxac")] < sub2$lcl[which(sub2$genotype==test.genotype)]){ 
                  emvec <- c(emvec,"res")
            }else{
                  emvec <- c(emvec,"nsd")
            }
      }
      names(emvec) <- unique(sub$dose)
      if(("res" %in% emvec) & ("sus" %in% emvec)){
            summary.matrix[test.pesticide,test.genotype] <- 0
      } else if("res" %in% emvec){
            summary.matrix[test.pesticide,test.genotype] <- 1
      } else if("sus" %in% emvec){
            summary.matrix[test.pesticide,test.genotype] <- -1
      } else {
            summary.matrix[test.pesticide,test.genotype] <- 0
      }
} 
abcb.hm <- data.frame(summary.matrix,stringsAsFactors = F)           

colnames(abcb.hm) <- multi.gsub(colnames(abcb.hm),find=c("wxac","mdr49ko","mdr50ko","mdr65ko"),c("Wxac","Mdr49KO","Mdr50KO","Mdr65KO"))
rownames(abcb.hm) <- sapply(rownames(abcb.hm),firstup)

abcb.hm$Wxac <- NULL

```


```{r Plot figure 1,fig.width=10}
par(mar = c(5, 12, 4, 2),family="serif")
color2D.matplot(abcb.hm, 
                show.values = FALSE,
                axes = FALSE,
                xlab = "",
                ylab = "",
                vcex = 2,
                vcol = "black",
                extremes = c("red4", "green4"), 
                show.legend = F,
                nslices=3)
axis(3, at = seq_len(ncol(abcb.hm)) - 0.5,
     labels = names(abcb.hm), tick = FALSE, cex.axis = 1.75, face="bold")
axis(2, at = seq_len(nrow(abcb.hm)) -0.5,
     labels = rev(rownames(abcb.hm)), tick = FALSE, las = 1, cex.axis = 1.5,face="bold")
```


```{r Import data for figure S3}
m65.metabolites <- read.csv("FigS3_M65_Metabolites.csv")
```

```{r Format data for figure S3}
m65.metabolites.corrected <- abbots.correction.table(m65.metabolites,control="wxac",write=F)
annot <- data.frame(x = rep(-Inf, length(unique(m65.metabolites.corrected$genotype))), y = rep(Inf, length(unique(m65.metabolites.corrected$genotype))), pesticide=as.character(unique(m65.metabolites.corrected$pesticide)),labs=LETTERS[1:length(unique(m65.metabolites.corrected$genotype))])
m65.metabolites.corrected$genotype <- factor(m65.metabolites.corrected$genotype,levels=c("wxac","mdr65ko"))
m65.metabolites.corrected$pesticide <- factor(m65.metabolites.corrected$pesticide,levels=c("imi-olefin","imi-5-oh"))
```

```{r Plot figure S3}
gp <- ggplot(m65.metabolites.corrected,aes(x=dose,y=corrected.survival,fill=genotype)) 
      gp <- gp+facet_grid(.~pesticide,scales="free_x",labeller = label_parsed)
      gp <- gp+geom_bar(position=position_dodge(width=.5),stat="identity",colour="black",width=.5)
      gp <- gp+geom_errorbar(aes(ymin=corrected.survival+conf.int,ymax=corrected.survival-conf.int),
                             position=position_dodge(width=.5),stat="identity",width=.5) 
      gp <- gp+ylab("Corrected Survival \n")
      gp <- gp+xlab("\n Concentration of Nitenpyram (ppm)")
      gp <- gp+scale_fill_manual(values=c("grey25","skyblue")) 
      gp <- gp+scale_y_continuous(breaks=c(seq(0,1,.2)))
      gp <- gp+geom_text(size=7,data=annot,aes(x, y, label=labs,fontface="bold",family="serif"), inherit.aes=FALSE,vjust=1.2,hjust=-.2)
      gp <- gp+theme_bw(base_size=14,base_family="serif") 
      gp <- gp+theme(text=element_text(face="bold",family="serif"),panel.grid=element_blank(),axis.ticks.x=element_line(),
                     panel.border=element_rect(colour="black",fill=NA),strip.text=element_text(size=18,face="bold"),
                     strip.background=element_rect("grey95"),
                     axis.title=element_text(size=18),axis.text.x=element_text(angle=30,hjust=1))
      
print(gp) 
```


```{r Read Data in for figure 2}
rr <- read.csv("Fig2_RR_Summary.csv") %>%
      arrange(Lethal.Dose.Ratio)
```

```{r Format data for figure 2}
rr$Insecticide <-  factor(rr$Insecticide,levels=rr$Insecticide)
col <- colorRampPalette(c("grey90","grey25"))(length(rr$Insecticide))
```

```{r Plot figure 2}
gp <- ggplot(data=rr,aes(x=Insecticide,y=Lethal.Dose.Ratio))
      gp <- gp+geom_bar(position="dodge",stat="identity",colour="black",fill="grey50")
      gp <- gp+geom_hline(yintercept = 1,linetype=2)
      gp <- gp+geom_errorbar(aes(ymin=LCL,ymax=UCL),width=.7,size=.5)
      gp <- gp+labs(y="Susceptibility Ratio\n (Mdr65KO/Wxac)\n",x="\nPesticide")
      gp <- gp+scale_y_continuous(expand = c(0, 0),limits=c(0,9),
                                  breaks=c(seq(0,8,2)),labels=c(seq(0,8,2)))
      gp <- gp+geom_text(size=4,label="N.S",aes(1, 1.75,family="serif"))
      gp <- gp+theme_bw(base_size=14,base_family="serif")
      gp <- gp+theme(text=element_text(face="bold",family="serif"),panel.grid=element_blank(),axis.ticks.x=element_line(),
                                 panel.border=element_rect(colour="black",fill=NA),
                                 axis.title=element_text(size=18),axis.text.x=element_text(angle=30,hjust=1),legend.position = "none")
      
      print(gp)
```


```{r Import data for Figure 3}
verapamil <- read.csv("Fig3_Verapamil_Synergism.csv")
```


```{r Format data for Figure 3}
ver.corrected <- lapply(split(verapamil,verapamil$Synergist),abbots.correction.table,control="wxac",write=F)
ver.corrected[["None"]]$synergist <- "None"
ver.corrected[["Verapamil"]]$synergist <- "Verapamil"
ver.summary <- rbindlist(ver.corrected) %>% data.frame()
ver.summary$genotype <- factor(ver.summary$genotype,levels=c("wxac","Mdr65KO"))

for(i in 1:nrow(ver.summary)){
      if(is.na(ver.summary[i,"genotype"])){ ver.summary[i,"genotype"] <- "Mdr65KO"}
}
annot <- data.frame(x = rep(-Inf, length(unique(ver.summary$genotype))), y = rep(Inf, length(unique(ver.summary$genotype))), genotype=as.character(unique(ver.summary$genotype)),labs=LETTERS[1:length(unique(ver.summary$genotype))])

```

```{r}
gp <- ggplot(ver.summary,aes(x=dose,y=corrected.survival,fill=synergist,group=interaction(dose,synergist))) 
      gp <- gp+facet_grid(.~genotype,scales="free_x",labeller = label_parsed)
      gp <- gp+geom_bar(position=position_dodge(width=.5),stat="identity",colour="black",width=.5)
      gp <- gp+geom_errorbar(aes(ymin=corrected.survival+conf.int,ymax=corrected.survival-conf.int),
                             position=position_dodge(width=.5),stat="identity",width=.5) 
      gp <- gp+ylab("Corrected Survival \n")
      gp <- gp+xlab("\n Concentration of Nitenpyram (ppm)")
      gp <- gp+scale_fill_manual(values=c("grey50","red")) 
      gp <- gp+scale_y_continuous(breaks=c(seq(0,1,.2)))
      gp <- gp+geom_text(size=7,data=annot,aes(x, y, label=labs,fontface="bold",family="serif"), inherit.aes=FALSE,vjust=1.2,hjust=-.2)
      gp <- gp+theme_bw(base_size=14,base_family="serif") 
      gp <- gp+theme(text=element_text(face="bold",family="serif"),panel.grid=element_blank(),axis.ticks.x=element_line(),
                     panel.border=element_rect(colour="black",fill=NA),strip.text=element_text(size=18,face="bold"),
                     strip.background=element_rect("grey95"),
                     axis.title=element_text(size=18),axis.text.x=element_text(angle=30,hjust=1))
      
print(gp) 
```

## Figure 4

```{r Import data for figure 4}

def <- read.csv("Fig4_Mdr65_Deficiency.csv")

```

```{r Format data for figure 4}
def$Genotype <- factor(def$Genotype,levels=c("Wxac x w1118","Mdr65KO x w1118", "Mdr65KO x Excel_DF", "Mdr65KO x Bloom_DF", "Mdr65KO x Mdr65KO"))
col <- colorRampPalette(c("grey90","grey25"))
```

```{r Plot figure 4}
 gp <- ggplot(def,aes(x=Genotype,y=Alive,fill=Genotype))
      gp <- gp+stat_summary(geom="errorbar",colour='black',size=1,width=.4,fun.data="mean_cl_normal")
      gp <- gp+stat_summary(fun.y="mean",fun.ymax="mean",fun.ymin="mean", geom="crossbar", width=.4,colour='black',size=.8)
      gp <- gp+geom_dotplot(binaxis='y',stackdir="center",binwidth=.75)
      gp <- gp+scale_fill_manual(values=c("black","grey50","white","white","white"))
      gp <- gp+ylab("Percent Survival\n")
      gp <- gp+xlab("\nGenotype")
      gp <- gp+theme_bw(base_size=14,base_family="serif") 
      gp <- gp+theme(text=element_text(face="bold",family="serif"),panel.grid=element_blank(),axis.ticks.x=element_line(),
                        panel.border=element_rect(colour="black",fill=NA),strip.text=element_text(size=18,face="bold"),
                        strip.background=element_rect("grey95"),
                        axis.title=element_text(size=18),axis.text.x=element_text(angle=30,hjust=1),legend.position = "none")
      
      print(gp)
```

```{r figure 4 stats}
mod <- with(def,aov(Alive~Genotype))
TukeyHSD(x=mod, which='Genotype', conf.level=0.95)$Genotype[,"p adj"]
```

```{r Import data for figure 5}
hplc <- read.csv("Fig5_HPLC_MS_Nitenpyram.csv",stringsAsFactors = F)
options(scipen=999)
```

```{r Format data for figure 5}
options(scipen=999)
Tissue <- c("Body","Head")
annot <- data.frame(x = rep(-Inf, length(unique(hplc$Tissue))), y = rep(Inf, length(unique(hplc$Tissue))), 
                    Tissue=as.character(unique(hplc$Tissue)),labs=rev(LETTERS[1:length(unique(hplc$Tissue))]))


hplc <- hplc[rev(order(hplc$Genotype)),]
hplc$Genotype <- factor(hplc$Genotype,levels=c("Wxac","Mdr65KO"))
hplc$Nitenpyram2 <- hplc$Nitenpyram/10000
```

```{r Plot figure 5}
 gp <- ggplot(hplc,aes(x=Genotype,y=Nitenpyram2,fill=Genotype))
      gp <- gp+facet_grid(Tissue~.,scales="free_y")
      gp <- gp+geom_dotplot(binaxis='y',stackdir="center",dotsize=2) 
      gp <- gp+scale_fill_manual(values=c("grey75","grey25"))
      gp <- gp+ylab(label="Quantity of Nitenpyram\n(Area Under Chromatogram/10,000)\n")
      gp <- gp+xlab("\nGenotype")
      gp <- gp+geom_text(size=7,data=annot,aes(x, y, label=labs,fontface="bold",family="serif"), inherit.aes=FALSE,vjust=1.2,hjust=-.2)
      gp <- gp+theme_bw(base_size=14,base_family="serif") 
      gp <- gp+theme(text=element_text(face="bold",family="serif"),panel.grid=element_blank(),axis.ticks.x=element_line(),
                     panel.border=element_rect(colour="black",fill=NA),strip.text=element_text(size=18,face="bold"),
                     strip.background=element_rect("grey95"),
                     axis.title=element_text(size=18),axis.text.x=element_text(angle=30,hjust=1),legend.position = "none")
      
      print(gp)
```


